## Implementing Williamson's analysis of income disparity to 1980 \~ 2018 Japan

Getting prefectural income per capita data from the Cabinet Office "県民経済計算" <https://www.esri.cao.go.jp/jp/sna/sonota/kenmin/kenmin_top.html>

```{r}
# ipc: income per capita
library(readxl)
clear_income_data <- function(file_path) {
  data_frame <- read_excel(file_path, sheet = "実数")
  colnames(data_frame) <- data_frame[1,] # rename the column names to gregoric years
  data_frame <- data_frame[3:50,] # get the prefecture portion out (get rid of other measures)
  prefecture_names <- as.matrix(data_frame[,2]) # change the row names to prefecture names
  data_frame <- data_frame[,5:ncol(data_frame)-1] # get rid of unnecessary columns (prefecture number and stuff)
  row.names(data_frame) <- prefecture_names
  return(data_frame)
}

ipc_2006_2018 <- clear_income_data("data/一人あたりの県民所得 2006~2018.xlsx")
ipc_1996_2009 <- clear_income_data("data/一人あたりの県民所得 1996~2009.xls")
ipc_1990_2003 <- clear_income_data("data/一人あたりの県民所得 1990~2003.xls")

prefecture_names <- rownames(ipc_1990_2003)

ipc_1990_2018 <- cbind(ipc_1990_2003[,1:6], ipc_1996_2009[,1:11], ipc_2006_2018)

row.names(ipc_1990_2018) <- prefecture_names
```

Getting Prefecture population from the Cabinet Office "県民経済計算" <https://www.esri.cao.go.jp/jp/sna/sonota/kenmin/kenmin_top.html>

```{r}
population_2006_2018 <- clear_income_data('data/総人口 2006~2018.xlsx')
population_2001_2014 <- clear_income_data('data/総人口 2001~2014.xls')
population_1990_2003 <- clear_income_data('data/総人口 1990~2003.xlsx')

population_1990_2018 <- cbind(population_1990_2003[,1:11], population_2001_2014[,1:5] ,population_2006_2018)

# deal with some character values
population_1990_2018 <- as.data.frame(lapply(population_1990_2018, function(x) as.numeric(as.character(x))))
population_1990_2018 <- t(population_1990_2018)

colnames(population_1990_2018) <- prefecture_names
population_1990_2018 <- as.data.frame(population_1990_2018)
test <- row.names(population_1990_2018) %>% map(function(x) as.numeric(substring(x, 2)))

population_1990_2018$year <- 1990:2018
```

Organizing GDP data
```{r}
gdp_data <- read_excel('data/GDP data of countries.xls', sheet = 'Data')
colnames(gdp_data) <- gdp_data[3,]
japan <- gdp_data[123,]
```

Calculating Williamson's inequality measure, which ought to measure the income disparity of the country

$$
\frac{\sqrt{\sum_{i}(y_i - \bar y)^2 * \frac{f_i}{n}}}{\bar y}\\
\bar y = \text{national incomeper capita}\\
y_i = \text{income per capita of the ith prefecture}\\
f_i = \text{population of the ith prefecture}\\
n = \text{population of Japan}\\
$$

Estimation model from Marius Brülhart Federica Sbergami (2008)

$$
g_{p} = \alpha y_{p} + \beta A_{p} + u_p\\

g_p = \text{average GDP growth of Japan over period p}\\
y_p = \text{log initial GDP of that period}\\
A_p = \text{primacy at the beginning of the period}\\
u_p = \text{error term}
$$

## Primacy
```{r}
shutoken <- c('東京都', '千葉県', '神奈川県', '埼玉県', '茨城県', '栃木県', '群馬県', '山梨県')
kanto <- c('東京都', '千葉県', '神奈川県', '埼玉県', '茨城県', '栃木県', '群馬県')

# Add a row for shutoken
population_1990_2018[,ncol(population_1990_2018) + 1] <- rowSums(population_1990_2018[,shutoken])
colnames(population_1990_2018)[ncol(population_1990_2018)] <- '首都圏'

# Add a row for kanto
population_1990_2018[,ncol(population_1990_2018) + 1]<- rowSums(population_1990_2018[,kanto])
colnames(population_1990_2018)[ncol(population_1990_2018)] <- '関東'

# primacy
primacy <- population_1990_2018[,'首都圏'] / population_1990_2018[,'全県計']
primacy <- cbind(primacy, population_1990_2018[,'東京都'] / population_1990_2018[,'全県計'])
primacy <- cbind(primacy, population_1990_2018[,'関東'] / population_1990_2018[,'全県計'])
primacy <- cbind(primacy, population_1990_2018[,'year'])

primacy <- data.frame(primacy)
colnames(primacy) <- c('Shutoken Primacy', 'Tokyo Primacy', 'Kanto Primacy', 'year')

primacy
library(ggplot2)
```

## Getting GDP from FRED
gdp in billions of chained 2015 yen from 1994 to 2021
```{r}
jpn_rgdp <- read_excel('data/JPNRGDPEXP.xls')

jpn_rgdp <- as.data.frame(cbind(1994:2021, colMeans(matrix(jpn_rgdp$JPNRGDPEXP, nrow=4))))

jpn_rgdp[,1]
colnames(jpn_rgdp) <- c('yaer', 'average real GDP of the year')
jpn_rgdp_xts <- xts(jpn_rgdp[,-1], order.by=as.Date(as.character(jpn_rgdp[,1]), "%Y"))
yearly_diff <- diff(jpn_rgdp_xts, lag = 1)
lagged <- jpn_rgdp_xts
index(lagged) <- index(lagged) %m+% years(1)
library(lubridate)
yearly_diff <- cbind(yearly_diff, lagged)
yearly_diff$growth <- yearly_diff$yearly_diff / yearly_diff$lagged

growth_rate <- na.omit(yearly_diff$growth)
log_gdp <- log(jpn_rgdp_xts)
```

